Explorations to Optimize the Parareal Algorithm \\to Solve ODEs in Given Applications

### 2014-12-09 paper.tex

Another obstacle to trying to paralleliize numerical methods is that the methods are inherently serial in time - in that evaluation of the next time $n+1$ depends on the previous values, say those at time $n-1$ and $n$.  This setup does not allow for parallelism.

@@

Show that this parareal method will approach, with large enough k (enough iterations) to approach the error of the fine method.  However, with too large of a k and a lower quality factor (ratio of fine to coarse), the time for the parareal could potentially take longer than the direct serial computation.

@@

The parareal algorithm, developed by Lions, Maday and Turinici in 2001, CITE is a general algorithm that allows parallelization in time.  It does so by using a cheaper (lower order, lesser resolution) approximation first, and then makes corrections in parallel.  The entire process is then iterative and tuned for a given amount of iterations.  Therefore, the algorithm is k repetitions of a serial coarse method updated by a finer method run in parallel for subsections.

\begin{algorithm}[t]
    \KwIn{Temporal discretization $t^n = t_0 + n \Delta t, \, n = 1,2,\ldots,N$}
    \KwIn{Coarse scheme $g_{\Delta t}$}
    \KwIn{Finer	 scheme $g_{\textnormal{fine}}$}
    Compute $u_1^{n+1} = g_{\Delta t}(t^n, u^n)$\;
    Compute the corrections $\delta g^n(u_1^n) = g_{\textnormal{fine}}(t^n,
    u^n) - g_{\Delta t}(t^n, u^n)$ in parallel\;
    Add the prediction and correction terms as $u_2^{n+1} = g_{\Delta
    t}(t^n, u_2^n) + \delta g^n(u_1^n)$\;
    Repeat steps 2 and 3, incrementing the iteration label and using $u_{k+1}^0
    = u^0$ as the initial condition\
 \caption{Parareal}
 \label{alg:parareal}
\end{algorithm}

\subsection{Visualization of Diagram}
bksim TODO

@@

\subsection{Explored Tests}
We wished to explore certain tests that would span the space of possible parameters to the parareal as well as possible applications in different ODE systems.  Our baseline was looking to compare

@@

\subsection{Comparison to Serial with $g_{fine}$ as \\a Forward Euler with
Smaller $\Delta t$}

@@

\subsection{Comparison to Serial with $g_{fine}$ as \\Higher Order Methods}

@@

\subsection{Comparison to a Forward Euler \\Parallelized by Space}

@@

\subsection{Comparison to C++ Implementations}
	The numpy environment combined with the required structure of MPI required certain type/data structure conversions which will take time.not not cunot cu

@@

ashi TODO by taking some numbers and following %https://www.sharcnet.ca/help/index.php/Measuring_Parallel_Scaling_Performance

@@

We see that (HOPE IT WORKS WELL OR ELSE...)
Most of the further work would be in optimization.  We are trying to measure performance in python, which is not the ideal benchmarking language for efficiency and speedups.  A port over to C++ using the MPI libraries in C++ rather than the mpi4py libraries in python would be ideal.  Other optimizations mentioned above coud also be implemented.  To further explore the parareal algorithm would involve many test cases and to see how the parareal algorithm compares to other approaches - since the actual time and iterations necessary depend on the system of ODEs to solve.

On the other hand, paralleism by space already requires a very problem-specific design.  The strong scaling efficiency of parallelism by space is high and is conceptually embarrisingly parallel and easier to create.  However, the limitations are in the requirement for the code to be designed for the problem - how to divide by space so as to allow for the maximum number of even divisions.  The lack of generalism is not as beautiful as the parareal algorithm but may lend itself to faster speedups.

All in all, we have implemented and begun to explore some techniques of looking for strong and weak scaling efficiencies for solving ODEs.  The parareal algorithm is beautiful in its theoretical advantages - in terms of stability, error and efficiency.  However, the method, as a very generalized method, requires deeper analysis for the specific problem.  The main variable will be in the difference in the coarse and fine methods (be it lower and higher order or a higher and lower resolution for the same method).  The tradeoffs that must be computed and optimized for would be problem specific.

@@

Bibtex Citations for the Slides, and for %https://www.sharcnet.ca/help/index.php/Measuring_Parallel_Scaling_Performance

>>>>

Another obstacle to trying to paralleliize numerical methods is that the methods
are inherently serial in time---in that evaluation of the next time $n+1$
depends on the previous values, say those at time $n-1$ and $n$. This setup
does not allow for parallelism.

@@

Show that this parareal method will approach, with large enough $k$ (enough
iterations) to approach the error of the fine method. However, with too large
of a k and a lower quality factor (ratio of fine to coarse), the time for the
parareal could potentially take longer than the direct serial computation.

@@

The parareal algorithm, developed by Lions, Maday and Turinici in 2001, CITE is
a general algorithm that allows parallelization in time.  It does so by using a
cheaper (lower order, lesser resolution) approximation first, and then makes
corrections in parallel. The entire process is then iterative and tuned for a
given amount of iterations. Therefore, the algorithm is $k$ repetitions of a
serial coarse method updated by a finer method run in parallel for subsections.

\begin{algorithm}[t]
    \KwIn{Temporal discretization $t^n = t_0 + n \Delta t, \, n = 1,2,\ldots,N$}
    \KwIn{Coarse scheme $g_{\Delta t}$}
    \KwIn{Finer	 scheme $g_{\textnormal{fine}}$}
    Compute $u^1_{n+1} = g_{\Delta t}(t_n, u^1_n)$\;
    Compute the corrections $\delta g_n(u^1_n) = g_{\textnormal{fine}}(t_n,
    u^1_n) - g_{\Delta t}(t_n, u^1_n)$ in parallel\;
    Add the prediction and correction terms as $u^2_{n+1} = g_{\Delta t}(t_n,
    u^2_n) + \delta g_n(u^1_n)$\;
    Repeat steps 2 and 3, incrementing the iteration label and using $u^{k+1}_0
    = u^1_0$ as the initial condition\;
 \caption{Parareal}
 \label{alg:parareal}
\end{algorithm}

\subsection{Visualization Diagram}
bksim TODO

@@

\subsection{Explored Tests}

We wished to explore certain tests that would span the space of possible
parameters to the parareal as well as possible applications in different ODE
systems. Our baseline was looking to compare

@@

\subsection{Comparison to Serial with $g_{fine}$ as a Forward Euler with
Smaller $\Delta t$}

\subsection{Comparison to Serial with $g_{fine}$ as Higher Order Methods}

@@

\subsection{Comparison to Serial with $g_{fine}$ as Higher Order Methods}

@@

\subsection{Comparison to a Forward Euler Parallelized by Space}

@@

\subsection{Comparison to C++ Implementations}

The numpy environment combined with the required structure of MPI required
certain type/data structure conversions which will take time.not not cunot
cu

@@

ashi TODO by taking some numbers and following
\url{https://www.sharcnet.ca/help/index.php/Measuring_Parallel_Scaling_Performance}

@@

We see that (HOPE IT WORKS WELL OR ELSE...) Most of the further work would be in
optimization. We are trying to measure performance in Python, which is not the
ideal benchmarking language for efficiency and speedups. A port over to C++
using the MPI libraries in C++ rather than the mpi4py libraries in python would
be ideal. Other optimizations mentioned above could also be implemented. To
further explore the parareal algorithm would involve many test cases and to see
how the parareal algorithm compares to other approaches---since the actual time
and iterations necessary depend on the system of ODEs to solve.

On the other hand, paralleism by space already requires a very problem-specific
design. The strong scaling efficiency of parallelism by space is high and is
conceptually embarrisingly parallel and easier to create. However, the
limitations are in the requirement for the code to be designed for the
problem---how to divide by space so as to allow for the maximum number of even
divisions. The lack of generalism is not as beautiful as the parareal algorithm
but may lend itself to faster speedups.

All in all, we have implemented and begun to explore some techniques of looking
for strong and weak scaling efficiencies for solving ODEs. The parareal
algorithm is beautiful in its theoretical advantages---in terms of stability,
error and efficiency. However, the method, as a very generalized method,
requires deeper analysis for the specific problem. The main variable will be in
the difference in the coarse and fine methods (be it lower and higher order or a
higher and lower resolution for the same method). The tradeoffs that must be
computed and optimized for would be problem specific.

@@

Bibtex Citations for the Slides, and for \url{https://www.sharcnet.ca/help/index.php/Measuring_Parallel_Scaling_Performance}
